Parsing mPFC: A Joint Meta-Analytic and Graph Theoretic Approach

### 2018-10-19 Manuscript.Rmd
Many updates, mainly to keep figures at the bottom

---
title: 'Parsing mPFC: A Joint Meta-Analytic and Graph Theoretic Approach'
author: "Claudio Toro-Serey^1^ & Joseph T. McGuire^1^"
output:
  pdf_document: default
  html_document: default
  word_document: 
    reference_docx: Style_reference.docx
bibliography: mPFC_paper.bib
header-includes:
    - \usepackage{setspace}\doublespacing
---

@@

# Let's compare stuff
testPCC <- RI_PCC[lower.tri(RI_PCC)]
testPFC <- RI_PFC[lower.tri(RI_PFC)]
Overall_PCCvsPFC_permTest <- permute(testPCC, testPFC, statType = mean, paired = T)
Overall_PCCvsPFC_ES <- cohenD(testPCC, testPFC) # caveat: maybe not ideal for bounded values, like RI. BUT VI yields similar results.
```

@@

``` {r Partition comparison: halves, echo=FALSE, message = FALSE, cache = TRUE}
# Select 10 subjects from the sample to display
selectSubj <- seq(1,20)

# Load
setwd('./Summary_halves/')
temp <- list.files(pattern = "*_finalSummary.csv")
subjList_halves <- sapply(temp, substring, first=1, last=9)
Summaries_halves <- lapply(temp, read.csv)
Summaries_halves_plot <- Summaries_halves[selectSubj]
subjList_halves_plot <- subjList_halves[selectSubj]

temp <- list.files(pattern = "*_corHalves.csv")
Correlation_halves <- lapply(temp, read.csv, header = F)
Correlation_halves_plot <- Correlation_halves[selectSubj]

# Compare all halves with RI
RI_all_between_halves <- comparePartitions(Data = Summaries_halves_plot, Index = "RI", nSubjects = length(Summaries_halves_plot), subjNames = subjList_halves_plot)

# Compare PCC and PFC among all subjects
tempComm_PCC_halves_plot <- lapply(Summaries_halves_plot, "[", i = posteriorIndx, j =)
tempComm_PFC_halves_plot <- lapply(Summaries_halves_plot, "[", i = !posteriorIndx, j =)
RI_PFC_halves <- comparePartitions(Data = tempComm_PFC_halves_plot, Index = "RI", nSubjects = length(Summaries_halves_plot), subjNames = subjList_halves_plot)
RI_PCC_halves <- comparePartitions(Data = tempComm_PCC_halves_plot, Index = "RI", nSubjects = length(Summaries_halves_plot), subjNames = subjList_halves_plot)
```

``` {r Partition comparison: halves ratios, echo=FALSE, message = FALSE, cache = TRUE}
## Now compare all possible pairs of subjects
# Compare all halves with RI
RI_all_between_halves <- comparePartitions(Data = Summaries_halves, Index = "RI", nSubjects = length(Summaries_halves), subjNames = subjList_halves)

@@

``` {r Partition comparison: session similarity matrices, echo=FALSE, message = FALSE, cache = TRUE}
# Select 10 subjects from the sample to display
selectSubj <- seq(1,40)

setwd('./Summary_sess/')
temp <- list.files()
subjList_sess <- sapply(temp, substring, first=1, last=9)
Summaries_sess <- lapply(temp, read.csv)
subjList_sess_plot <- subjList_sess[selectSubj]
Summaries_sess_plot <- Summaries_sess[selectSubj]

# Across subjects (just for plotting)
RI_all_between_sess <- comparePartitions(Data = Summaries_sess_plot, Index = "RI", nSubjects = length(Summaries_sess_plot), subjNames = subjList_sess_plot)
tempComm_PCC_sess <- lapply(Summaries_sess_plot, "[", i = posteriorIndx, j =)
tempComm_PFC_sess <- lapply(Summaries_sess_plot, "[", i = !posteriorIndx, j =)
RI_PFC_sess <- comparePartitions(Data = tempComm_PFC_sess, Index = "RI", nSubjects = length(Summaries_sess_plot), subjNames = subjList_sess_plot)
RI_PCC_sess <- comparePartitions(Data = tempComm_PCC_sess, Index = "RI", nSubjects = length(Summaries_sess_plot), subjNames = subjList_sess_plot)
```

``` {r Partition comparison: session compare all, echo=FALSE, message = FALSE, cache = TRUE}
# Now run comparisons among all subjects
RI_all_between_sess <- comparePartitions(Data = Summaries_sess, Index = "RI", nSubjects = length(Summaries_sess), subjNames = subjList_sess)
tempComm_PCC_sess <- lapply(Summaries_sess, "[", i = posteriorIndx, j =)
tempComm_PFC_sess <- lapply(Summaries_sess, "[", i = !posteriorIndx, j =)
RI_PFC_sess <- comparePartitions(Data = tempComm_PFC_sess, Index = "RI", nSubjects = length(Summaries_sess), subjNames = subjList_sess)
RI_PCC_sess <- comparePartitions(Data = tempComm_PCC_sess, Index = "RI", nSubjects = length(Summaries_sess), subjNames = subjList_sess)

@@

It will also be good to note that there is concurrent debate on the ground truth of community detection. In other fields, identifying a community structure is not necessarily associated with the true organization of the network. What happens is that you might find a tangential grouping that gives you information different from your original goal [@Clauset2004]

@@

Add a portion showing that summing up the binarized maps produces a map similar to the one produced by Margulies et al. and others.

@@

@@

``` {r Partition comparison: halves (Fig. 5), fig.align="center", out.width = "60%", fig.cap = "Figure 5. Similarity matrices for 10 example participants showing pattern agreement across days for the whole partition (left), and for PCC and mPFC separately (right). These figures show high agreement within (but not between) subjects, with mPFC again showing significant variability.", echo=FALSE, message = FALSE, cache = TRUE}
# plots
par(mfrow=c(1,2))
corrplot(RI_all_between_halves, 
         is.corr = F, 
         method = "color",
         addgrid.col = "grey90",
         tl.pos = "n", 
         tl.col = "black",
         tl.cex = 0.1,
         cl.pos = "n",
         cl.lim = c(0,1),
         cl.length = 3,
         mar = c(1.1,1.1,1.1,1.1))
# segments(0.5,0.5,0.5,20.5, lwd = 2)
# segments(0.5,0.5,20.5,0.5, lwd = 2)
# segments(0.5,20.5,20.5,20.5, lwd = 2)
# segments(20.5,0.5,20.5,20.5, lwd = 2)
mtext(expression(bold("Subjects (Day 1 and Day 2)")), line = 1)
mtext(expression(bold("Subjects (Day 1 and Day 2)")), side = 2, line = 3)
corrplot(RI_PCC_halves, 
         is.corr = F,     
         method = "color",
         type = "lower", 
         tl.pos = "n", 
         cl.lim = c(0,1),
         cl.pos = "r",
         addgrid.col = "grey90",
         cl.length = 3)
corrplot(RI_PFC_halves, 
         is.corr = F, 
         method = "color",
         type = "upper", 
         tl.pos = "n", 
         cl.pos = "n", 
         addgrid.col = "grey90",
         add = T)
mtext(expression(bold("Subjects (Day 1 and Day 2)")), line = 1)
mtext(expression(bold("mPFC")), line = -0.15, at = par("usr")[1]+0.8*diff(par("usr")[1:2]), col = "steelblue4")
mtext(expression(bold("PCC")), side = 1, line = 0.6, at = par("usr")[1]+0.09*diff(par("usr")[1:2]), col = "steelblue4")
```

@@

``` {r Partition comparison: session similarity matrices (Fig. 7), fig.align="center", out.width = "60%", fig.cap = "Figure 7. Similarity matrices showing pattern agreement across sessions for the whole partition (left), and for PCC and mPFC separately (right) (10 example subjects). The pattern from the daily analysis is preserved, with marked individuality for mPFC, but not PCC.", echo=FALSE, message = FALSE, cache = TRUE}
# plot
par(mfrow=c(1,2))
corrplot(RI_all_between_sess, 
         is.corr = F, 
         method = "color", 
         tl.pos = "n", 
         tl.cex = 0.5,
         tl.col = "black", 
         cl.lim = c(-0.03,1),
         cl.pos = "n",
         cl.length = 3,
         mar = c(1.1,1.1,1.1,1.1), 
         addgrid.col = "grey90")
# segments(0.5,0.5,0.5,40.5, lwd = 2)
# segments(0.5,0.5,40.5,0.5, lwd = 2)
# segments(0.5,40.5,40.5,40.5, lwd = 2)
# segments(40.5,0.5,40.5,40.5, lwd = 2)
mtext(expression(bold("Subjects (Session)")), line = 1)
mtext(expression(bold("Subjects (Session)")), side = 2, line = 3)
corrplot(RI_PCC_sess, 
         is.corr = F, 
         method = "color", 
         type = "lower", 
         tl.pos = "n", 
         tl.cex = 0.5,
         tl.col = "black", 
         cl.lim = c(0,1),
         cl.length = 3,
         cl.pos = "r", 
         addgrid.col = "grey90")
# segments(0.5,0.5,0.5,20.5, lwd = 2)
# segments(0.5,0.5,20.5,0.5, lwd = 2)
# segments(0.5,20.5,20.5,20.5, lwd = 2)
# segments(20.5,0.5,20.5,20.5, lwd = 2)
corrplot(RI_PFC_sess, 
         is.corr = F, 
         method = "color", 
         type = "upper", 
         tl.pos = "n", 
         tl.col = "black", 
         cl.pos = "n", 
         addgrid.col = "grey90",
         add = T)
mtext(expression(bold("Subjects (Session)")), line = 1)
mtext(expression(bold("mPFC")), line = 0, at = par("usr")[1]+0.8*diff(par("usr")[1:2]), col = "steelblue4")
mtext(expression(bold("PCC")), side = 1, line = 1, at = par("usr")[1]+0.09*diff(par("usr")[1:2]), col = "steelblue4")
```

>>>>

---
title: 'Parsing mPFC: A Joint Meta-Analytic and Graph Theoretic Approach'
author: "Claudio Toro-Serey^1^ & Joseph T. McGuire^1^"
output:
  word_document: 
    reference_docx: Style_reference.docx
  pdf_document: default
  html_document: default
bibliography: mPFC_paper.bib
header-includes:
    - \usepackage{setspace}\doublespacing
---

@@

# Let's compare stuff
testPCC <- RI_PCC[lower.tri(RI_PCC)]
testPFC <- RI_PFC[lower.tri(RI_PFC)]
Overall_PCCvsPFC_permTest <- permute(testPCC, testPFC, statType = mean, paired = T)
Overall_PCCvsPFC_ES <- cohenD(testPCC, testPFC) # caveat: maybe not ideal for bounded values, like RI. BUT VI yields similar results.

# To get the rate of stability per vertex across subjects
FB <- do.call(cbind, lapply(Summaries, "[[", "FiedlerBinary"))
FBS <- rowMeans(FB)
FBS[FBS < 0.5] <- 1 - FBS[FBS < 0.5]

# Now let's test something: 
# Traditionally, a vertex-wise test would be carried across subjects.
# If these traditional analyses miss out on these idiosyncrasies, you should get unsignificant results.
# So we did a prop.test per vertex, and corrected with FDR to mimic this situation.
All_proptests <- apply(FB, 1, function(x) {prop.test(sum(x), nSubj)})
Pvals_corrected <- p.adjust(sapply(All_proptests, "[[", "p.value"), method = "BY")
Chisquares <- sapply(All_proptests, "[[", "statistic")
```

@@

``` {r Partition comparison: halves, echo=FALSE, message = FALSE, cache = TRUE}
# Load data
setwd('./Summary_halves/')
temp <- list.files(pattern = "*_finalSummary.csv")
subjList_halves <- sapply(temp, substring, first=1, last=9)
Summaries_halves <- lapply(temp, read.csv)

# Select subsection of subjects for plotting the similarity matrices
selectSubj <- seq(1,20)
Summaries_halves_plot <- Summaries_halves[selectSubj]
subjList_halves_plot <- subjList_halves[selectSubj]

# Load the correlation-based maps (this might be ok to move to the cor vs comm chunk)
temp <- list.files(pattern = "*_corHalves.csv")
Correlation_halves <- lapply(temp, read.csv, header = F)
Correlation_halves_plot <- Correlation_halves[selectSubj]

# Compare all halves with RI (plot only)
RI_all_between_halves_plot <- comparePartitions(Data = Summaries_halves_plot, Index = "RI", nSubjects = length(Summaries_halves_plot), subjNames = subjList_halves_plot)

# Compare PCC and PFC among all subjects (plot only)
tempComm_PCC_halves_plot <- lapply(Summaries_halves_plot, "[", i = posteriorIndx, j =)
tempComm_PFC_halves_plot <- lapply(Summaries_halves_plot, "[", i = !posteriorIndx, j =)
RI_PFC_halves_plot <- comparePartitions(Data = tempComm_PFC_halves_plot, Index = "RI", nSubjects = length(Summaries_halves_plot), subjNames = subjList_halves_plot)
RI_PCC_halves_plot <- comparePartitions(Data = tempComm_PCC_halves_plot, Index = "RI", nSubjects = length(Summaries_halves_plot), subjNames = subjList_halves_plot)
```

``` {r Partition comparison: halves ratios, echo=FALSE, message = FALSE, cache = TRUE}
## Now compare all possible pairs of subjects (for actual reporting)
# Compare all halves with RI
RI_all_between_halves <- comparePartitions(Data = Summaries_halves, Index = "RI", nSubjects = length(Summaries_halves), subjNames = subjList_halves)

@@

``` {r Partition comparison: session similarity matrices, echo=FALSE, message = FALSE, cache = TRUE}
# Load data
setwd('./Summary_sess/')
temp <- list.files()
subjList_sess <- sapply(temp, substring, first=1, last=9)
Summaries_sess <- lapply(temp, read.csv)

# Select 10 subjects from the sample to plot
selectSubj <- seq(1,40)
subjList_sess_plot <- subjList_sess[selectSubj]
Summaries_sess_plot <- Summaries_sess[selectSubj]

# Across subjects (just for plotting)
RI_all_between_sess_plot <- comparePartitions(Data = Summaries_sess_plot, Index = "RI", nSubjects = length(Summaries_sess_plot), subjNames = subjList_sess_plot)
tempComm_PCC_sess <- lapply(Summaries_sess_plot, "[", i = posteriorIndx, j =)
tempComm_PFC_sess <- lapply(Summaries_sess_plot, "[", i = !posteriorIndx, j =)
RI_PFC_sess_plot <- comparePartitions(Data = tempComm_PFC_sess, Index = "RI", nSubjects = length(Summaries_sess_plot), subjNames = subjList_sess_plot)
RI_PCC_sess_plot <- comparePartitions(Data = tempComm_PCC_sess, Index = "RI", nSubjects = length(Summaries_sess_plot), subjNames = subjList_sess_plot)
```

``` {r Partition comparison: session compare all, echo=FALSE, message = FALSE, cache = TRUE}
# Now run comparisons among all subjects (for actual reporting)
RI_all_between_sess <- comparePartitions(Data = Summaries_sess, Index = "RI", nSubjects = length(Summaries_sess), subjNames = subjList_sess)
tempComm_PCC_sess <- lapply(Summaries_sess, "[", i = posteriorIndx, j =)
tempComm_PFC_sess <- lapply(Summaries_sess, "[", i = !posteriorIndx, j =)
RI_PFC_sess <- comparePartitions(Data = tempComm_PFC_sess, Index = "RI", nSubjects = length(Summaries_sess), subjNames = subjList_sess)
RI_PCC_sess <- comparePartitions(Data = tempComm_PCC_sess, Index = "RI", nSubjects = length(Summaries_sess), subjNames = subjList_sess)

@@

It will also be good to note that there is concurrent debate on the ground truth of community detection. In other fields, identifying a community structure is not necessarily associated with the true organization of the network. What happens is that you might find a tangential grouping that gives you information different from your original goal [@Clauset2004]. In our case, however, averaging everyone's partitions produces organizational maps that resemble those from previous work (e.g. Margulies).

@@

@@

So what's the advantage of this work over previous ones? Gratton, Gordon, and Kong have already found some degree of subject stability and intersubject variability. Ours complements these findings by looking at the vertex level, not just mean parcel activity (this made sense for them, as they were looking at the whole brain). Also, most of these analyses rely on a small set of oversampled subjects (i.e. MSC). Here we extend this to 100, giving credit to individualize examinations even without oversampling.

@@

``` {r Partition comparison: halves (Fig. 5), fig.align="center", out.width = "60%", fig.cap = "Figure 5. Similarity matrices for 10 example participants showing pattern agreement across days for the whole partition (left), and for PCC and mPFC separately (right). These figures show high agreement within (but not between) subjects, with mPFC again showing significant variability.", echo=FALSE, message = FALSE, cache = TRUE}
# plots
par(mfrow=c(1,2))
corrplot(RI_all_between_halves_plot, 
         is.corr = F, 
         method = "color",
         addgrid.col = "grey90",
         tl.pos = "n", 
         tl.col = "black",
         tl.cex = 0.1,
         cl.pos = "n",
         cl.lim = c(0,1),
         cl.length = 3,
         mar = c(1.1,1.1,1.1,1.1))
# segments(0.5,0.5,0.5,20.5, lwd = 2)
# segments(0.5,0.5,20.5,0.5, lwd = 2)
# segments(0.5,20.5,20.5,20.5, lwd = 2)
# segments(20.5,0.5,20.5,20.5, lwd = 2)
mtext(expression(bold("Subjects (Day 1 and Day 2)")), line = 1)
mtext(expression(bold("Subjects (Day 1 and Day 2)")), side = 2, line = 3)
corrplot(RI_PCC_halves_plot, 
         is.corr = F,     
         method = "color",
         type = "lower", 
         tl.pos = "n", 
         cl.lim = c(0,1),
         cl.pos = "r",
         addgrid.col = "grey90",
         cl.length = 3)
corrplot(RI_PFC_halves_plot, 
         is.corr = F, 
         method = "color",
         type = "upper", 
         tl.pos = "n", 
         cl.pos = "n", 
         addgrid.col = "grey90",
         add = T)
mtext(expression(bold("Subjects (Day 1 and Day 2)")), line = 1)
mtext(expression(bold("mPFC")), line = -0.15, at = par("usr")[1]+0.8*diff(par("usr")[1:2]), col = "steelblue4")
mtext(expression(bold("PCC")), side = 1, line = 0.6, at = par("usr")[1]+0.09*diff(par("usr")[1:2]), col = "steelblue4")
```

@@

``` {r Partition comparison: session similarity matrices (Fig. 7), fig.align="center", out.width = "60%", fig.cap = "Figure 7. Similarity matrices showing pattern agreement across sessions for the whole partition (left), and for PCC and mPFC separately (right) (10 example subjects). The pattern from the daily analysis is preserved, with marked individuality for mPFC, but not PCC.", echo=FALSE, message = FALSE, cache = TRUE}
# plot
par(mfrow=c(1,2))
corrplot(RI_all_between_sess_plot, 
         is.corr = F, 
         method = "color", 
         tl.pos = "n", 
         tl.cex = 0.5,
         tl.col = "black", 
         cl.lim = c(-0.03,1),
         cl.pos = "n",
         cl.length = 3,
         mar = c(1.1,1.1,1.1,1.1), 
         addgrid.col = "grey90")
# segments(0.5,0.5,0.5,40.5, lwd = 2)
# segments(0.5,0.5,40.5,0.5, lwd = 2)
# segments(0.5,40.5,40.5,40.5, lwd = 2)
# segments(40.5,0.5,40.5,40.5, lwd = 2)
mtext(expression(bold("Subjects (Session)")), line = 1)
mtext(expression(bold("Subjects (Session)")), side = 2, line = 3)
corrplot(RI_PCC_sess_plot, 
         is.corr = F, 
         method = "color", 
         type = "lower", 
         tl.pos = "n", 
         tl.cex = 0.5,
         tl.col = "black", 
         cl.lim = c(0,1),
         cl.length = 3,
         cl.pos = "r", 
         addgrid.col = "grey90")
# segments(0.5,0.5,0.5,20.5, lwd = 2)
# segments(0.5,0.5,20.5,0.5, lwd = 2)
# segments(0.5,20.5,20.5,20.5, lwd = 2)
# segments(20.5,0.5,20.5,20.5, lwd = 2)
corrplot(RI_PFC_sess_plot, 
         is.corr = F, 
         method = "color", 
         type = "upper", 
         tl.pos = "n", 
         tl.col = "black", 
         cl.pos = "n", 
         addgrid.col = "grey90",
         add = T)
mtext(expression(bold("Subjects (Session)")), line = 1)
mtext(expression(bold("mPFC")), line = 0, at = par("usr")[1]+0.8*diff(par("usr")[1:2]), col = "steelblue4")
mtext(expression(bold("PCC")), side = 1, line = 1, at = par("usr")[1]+0.09*diff(par("usr")[1:2]), col = "steelblue4")
```
